---
title: "QFish Assignment"
author: "Jeremy Medina"
date: "2023-05-17"
output: html_document
---
### **1. Code used to complete data analysis**


#### **1.1** Loading of the data provided by QFish as well as the packages required for data analysis.
```{r setup, warning=FALSE, results='hide', message=FALSE}
library("tidyr")
library("tibble")
library("tidyverse")
library("ggplot2")
library("dplyr")
shark_data = read.csv("~/Desktop/github/mb5370/data/export (7).csv")
```
<br>

#### **1.2** Tidying of data to ensure correct variable types as well as to remove unwanted or unusable pieces of data for R.
```{r, results='hide'}
newsharkdata <- shark_data[-c(35,34),]
#Removing rows 34 and 35 as these include totals and 2023 which currently has incomplete data collection.

newsharkdata$CalendarYear <- as.numeric(as.character(newsharkdata$CalendarYear))
str(newsharkdata$CalendarYear)
#Changing the calendar year variable type to be numerical rather than characters.

newsharkdata$Total.Sharks.Caught <- as.numeric(as.character(newsharkdata$Total.Sharks.Caught))
str(newsharkdata$Total.Sharks.Caught)
#Changing the count of sharks discarded variable type to numerical.

```
<br>

#### **1.3** Code for creating figure 1.
```{r, results='hide', fig.keep='none'}
ggplot(newsharkdata, aes(x = CalendarYear, y = Total.Sharks.Caught)) + geom_line() + labs(title = "Discarded sharks caught by QLD fisheries 1990-2022", x = "Years", y ="Number of sharks discarded")+ theme(plot.title = element_text(hjust = 0.5)) + 
  annotate(geom = "text", x=2007, y=47630, label="2018 peak may be the result of improved management") + theme(panel.grid.major = element_blank(),
               panel.grid.minor = element_blank())
```

<br>

#### **1.4** Code to allow us to see the number of sharks discarded by year, I sorted by largest to find out that 2018 was the year that had the highest number of sharks discarded.
```{r,results='hide'}
twosharkdata <- data.frame(newsharkdata$CalendarYear, newsharkdata$Total.Sharks.Caught)
problemyears <- slice_max(twosharkdata, order_by = newsharkdata.Total.Sharks.Caught, n = 10)
view(twosharkdata)
```

<br>

### **2. Analysis**



#### The data I chose to work with was Queensland commercial and charter fisheries data provided by the Queensland Department of Agriculture and Fisheries (DAF) QFish datasets. The  data I was particularly interested in and analyzed involved the number of sharks discarded from 1990-2022 as reported by the DAF.  I was able to create a line graph to determine potential changes in the number of sharks discarded in the previously mentioned time period. I tidied the data by removing the totals provided for each year as well as removing the incomplete data for 2023. Furthermore I was required to change the year and discarded shark count variable types to numerical as they were loaded in as characters and integers respectively. 

<br>

#### After the data was tidied, I utilized ggplot to visualize the data and create figure 1. In addition, I sorted the years within 1990-2022 by the number of sharks discarded per year to determine the year with the highest number of discarded sharks.

<br>

```{r, echo=FALSE}
ggplot(newsharkdata, aes(x = CalendarYear, y = Total.Sharks.Caught)) + geom_line() + labs(title = "Discarded sharks caught by QLD fisheries 1990-2022", x = "Years", y ="Number of sharks discarded")+ theme(plot.title = element_text(hjust = 0.5)) + 
  annotate(geom = "text", x=2008, 47630, label="2018 peak may be due to improved management") + theme(panel.grid.major = element_blank(),
               panel.grid.minor = element_blank())
```

#### **Figure 1.** Line graph created from the discarded shark data collected by the Queensland Department of Agriculture and Fisheries from 1990-2022. The year in which the number of discarded sharks was highest was 2018.

<br>

### **3. Conclusion**



#### After the creation of figure 1., I completed an additional step in my analysis to determine the year in which the number of discarded sharks was the highest, resulting in determining that 2018 had the highest number of discarded sharks at 47,630. When researching this spike in the number of discarded sharks, I found that there was a shark assessment report completed in 2018 by the Australian Bureau of Agricultural and Resource Economics and Sciences, which introduced additional legislation and protections for shark stocks, improvements to catch and effort data collection, and the improved identification of caught sharks (DAFF, n.d.). Furthermore, it is believed that Australia's national Shark-plan 2 of 2018 has also contributed to improving the conservation and management of shark stocks in Australian waters (DAFF, n.d.), which may have also played a role in the 2018 spike noted in the line graph.

<br>

### **Literature Cited**

#### Australian Government: Department of agriculture, fisheries, and forestry. (n.d.). Sharks. Sharks - DAFF. https://www.agriculture.gov.au/agriculture-land/fisheries/environment/sharks#daff-page-main 

